
%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{May 9, 2021}
\chapter{Geometry TSTST}\label{tstst:geo}
%National camp Exam 2
%:'(
The Geometry Exam of The National Camp.
%Question
Problems are given below:

\begin{prob}\label{geo:1}
Let $\triangle ABC$ be a triangle inscribed in a circle $\omega$. $D, E$ are two points on the arc $BC$
of $\omega$ not containing $A$. Points $F, G$ lie on $BC$ such that
\[\angle BAF =\angle CAD %\mathrm{and} 
\text{ and } \angle BAG=\angle CAE\]
Prove that the two lines $DG$ and $EF$ meet on $\omega$.
\end{prob}

\begin{prob}\label{geo:2}
$ABC$ is a triangle where $\angle BAC = 90\dg$ . A line through the midpoint $D$ of $BC$ meets
$AB$ at $X$ and $AC$ at $Y$ . The point $P$ is taken on $XY$ such that $PD$ and $XY$ have the same midpoint
$M$. The perpendicular from $P$ to $BC$ meets $BC$ at $T$ .

Prove that AM bisects $\angle TAD$.
\end{prob}

\begin{prob}\label{geo:3}
Let $ABC$ be a triangle with $BC$ being the longest side. Let $O$ be the circumcenter
of $ABC$. $P$ is an arbitrary point on $BC$. The perpendicular bisector of $BP$ meet $AB$ at $Q$ and the
perpendicular bisector of $PC$ meet $AC$ at $R$. Prove that $AQOR$ is cyclic.
\end{prob}

\textit{After solving all of the three problems above, you may try this bonus problem. Beware though,
any work done on Problem 4 with incomplete solutions of the first three won’t receive any
points.}

\begin{prob}\label{geo:4}
Let $ABC$ be a triangle with circumcircle $(O)$. The midpoints of $BC,CA, AB$ are
$A', B', C'$ respectively. The medians $AA', BB' ,CC'$ cut the circumcircle $(O)$ at $A, A_1 ; B, B_1 ; C,C_1$
respectively. The line of tangency to $(O)$ at $A_1$ meets the perpendicular to $AO$ through $A ′$ at $X$.
Define $Y, Z$ similarly. Prove that $X,Y, Z$ are collinear.
\end{prob}



Full Solutions to these problems can be found at \autoref{sols:tstst}.